---
title: "Covid91 vaccine study Final2023"
author: "Ashley Cook"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "July 31, 2023" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
males = filter(FinalData, sex =="M")
females = filter(FinalData, sex =="F")
LGBTQ = filter(FinalData, LGBTQ =="gay")
Druggies = filter(FinalData, DrugUser =="yes")

```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

##Methods

In this Final, we will be looking at data from each group that have got the placebo and had a reaction from it as in getting sick; and, how many people got covid without taking the vaccine. The groups are Males, Females, LGBTQ and Druggies. We will be looking at data from barplots, numerical results, inferential results, and the effectiveness confidence interval. At the very end, I will explain which group was most effective from taking the vaccine for covid. 

## Males

```{r}
dd2 <- males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
This barplot shows how many males got effected by the vaccine or the placebo. The data shows that there is more males that took the vaccine but still ended up getting covid-19. Even some that took the placebo which was a smaller percentage than the vaccine, still got covid-19 as well. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=males)
rowPerc(table1)
colPerc(table1)
```
  
These results show the percentages of males who got covid-19 from taking the placebo and those who did not. Those who took the placebo which is 69% got covid-19, while only 30%  got covid-19 from taking the vaccine. So the vaccine does not prevent males getting covid-19 as well as taking the placebo.  

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
malesfish <- fisher.test(table1)
malesfish
```
  
For this chi-squared test, shows that the males who took the vaccine are more likely to get covid-19. The p-value for this shows a 4.631e-14, which is not a good number to see. The lower the p-value have the better results. Covid-19 does show that the illness can also be based off gender, since each body is different and reacts differently from covid-19. 

## Effectiveness Confidence Interval

```{r}
effect <- malesfish$p.value/(malesfish$p.value +1)
effectivenessguy = malesfish$estimate/(malesfish$estimate +1)*100
print(effectivenessguy)
```
```{r}
low=(malesfish$conf.int[1]/(malesfish$conf.int[1] +1))*100
hi=(malesfish$conf.int[2]/(malesfish$conf.int[2] +1))*100
options(digits=2)
```
  
For the odds ratio it shows that it is at a 69.15981, which is a high number that the odds of people taking the vaccine are getting effected by covid-19. 


## Females

```{r}
dd2 <- females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
This plot shows that the vaccine protected more females from covid-19 than the placebo. There was a higher percentage of females who took the placebo but still ended up getting infected. There was 60% of females who got infected by taking placebo and about 39% from the vaccine. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```
  
About 60 females got infected by covid-19 from taking the placebo while 40 females got infected by covid-19 from the vaccine. Out of 100 females, 50 on both sides were okay from taking the placebo and the vaccine. For the 2nd group, there were more people that did not get covid-19 from either the placebo or vaccine. 

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
femalesfish <- fisher.test(table1)
femalesfish
```
  
Since the p-value (3e-04) is less than the conventional significance level of .05, we can reject the null hypothesis and conclude that there is a significant association between the infection status and the treatment.

## Efectiveness Confidence Interval

```{r}
effect <- femalesfish$p.value/(femalesfish$p.value +1)
effectivenessfemale = femalesfish$estimate/(femalesfish$estimate +1)*100
print(effectivenessfemale)
```
```{r}
low=(femalesfish$conf.int[1]/(femalesfish$conf.int[1] +1))*100
hi=(femalesfish$conf.int[2]/(femalesfish$conf.int[2] +1))*100
options(digits=2)
```
  
For the females, they have an odds ratio of 60 while the males had 69. The males were more effected so far by covid-19 than the females. So by this, shows that there was a reaction between the placebo and the vaccine. 


## LGBTQ

```{r}
dd2 <- LGBTQ %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
By the looks of this chart, the LGBTQ community was more affected by taking the vaccine than the placebo. There was a percentage of around 15% of LGBTQ that took the placebo, while 85% took the vaccine and got covid-19. Than the second group, shows that 100 LGBTQ took the placebo and vaccine but did not get covid-19.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
  
Based on the numerical results, the vaccine took an effect on the LGBTQ community rather than the placebo took more an effect on the males and females. But, the LGBTQ community did have more of a higher percentage of people who were okay after taking the placebo and vaccine, than the other two groups. 

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
LGBTQfish <- fisher.test(table1)
LGBTQfish
```
  
This shows that there is a p-value of 7e-08. This indicates a highly significant association between the infection status and the treatment.This shows that the vaccine has a higher effectiveness rate that does not protect you from covid-19. 


## Effectiveness Confidence Interval

```{r}
effect <- LGBTQfish$p.value/(LGBTQfish$p.value +1)
effectivenessLGBTQ = LGBTQfish$estimate/(LGBTQfish$estimate +1)*100
print(effectivenessLGBTQ)
```

```{r}
low=(LGBTQfish$conf.int[1]/(LGBTQfish$conf.int[1] +1))*100
hi=(LGBTQfish$conf.int[2]/(LGBTQfish$conf.int[2] +1))*100
options(digits=2)
```
  
The odd ratio for the LGBTQ is 15. 

## Druggies

```{r}
dd2 <- Druggies %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
For the druggies, the vaccine effected them more than by taking the placebo to prevent covid-19. With the percentage of 81% of druggies who took the vaccine and 9% of druggies who took the placebo. More druggies were not effected by the placebo than the vaccine by like 2%.

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Druggies)
rowPerc(table1)
colPerc(table1)
```
  
More druggies did get effected by taking the vaccine and ended up getting covid-19 than the rest who took the placebo. group 2 had more druggies who did not get effected at all, than the first group. But in the first group, more people did not get covid-19 by taking the placebo than the vaccine. 



## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
Druggiesfish <- fisher.test(table1)
Druggiesfish
```
  
There is a p-value of 3e-09. This supports that there is strong evidence and effectiveness between the infected and treatment rates. 

## Effectiveness Confidence Interval

```{r}
effect <- Druggiesfish$p.value/(Druggiesfish$p.value +1)
effectivenessDruggie = Druggiesfish$estimate/(Druggiesfish$estimate +1)*100
print(effectivenessDruggie)
```

```{r}
low=(Druggiesfish$conf.int[1]/(Druggiesfish$conf.int[1] +1))*100
hi=(Druggiesfish$conf.int[2]/(Druggiesfish$conf.int[2] +1))*100
options(digits=2)
```

The odds ratio is a 11, which is the lowest out of all the groups. This shows that there was not a huge reaction between the two varuiables vaccine and placebo.

# Overall Results and Conclusions

Out of the 4 groups we looked at data from, the highest odd ratio was from the males with a 69. The females had the smallest p-value of 3e-04 which supports the null hypothesis and supports stronger evidence. The druggies got infected from covid-19 by taking the vaccine the most out of the 4 groups. The males got infected by covid-19 from taking the placebo the most out of the 4 groups. When it comes to the 2nd group of data from each group, the 2nd group did good when it came to getting infected from either the placebo or vaccine. 








