---
title: "Covid91 vaccine study Final2023"
author: "Ashley Cook"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "July 31, 2023" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
males = filter(FinalData, sex =="M")
females = filter(FinalData, sex =="F")
LGBTQ = filter(FinalData, LGBTQ =="gay")
Druggies = filter(FinalData, DrugUser =="yes")

```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

##Methods

In this Final, we will be looking at data from each group that have got the placebo and had a reaction from it as in getting sick; and, how many people got covid without taking the vaccine. The groups are Males, Females, LGBTQ and Druggies. We will be looking at data from barplots, numerical results, inferential results, and the effectiveness confidence interval. At the very end, I will explain which group was most effective from taking the vaccine for covid. 

## Males

```{r}
dd2 <- males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
This barplot shows how many males got effected by the vaccine or the placebo. The data shows that there is more males that took the vaccine but still ended up getting covid-19. Even some that took the placebo which was a smaller percentage than the vaccine, still got covid-19 as well. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=males)
rowPerc(table1)
colPerc(table1)
```
  
These results show the percentages of males who got covid-19 from taking the placebo and those who did not. Those who took the placebo which is 69% got covid-19, while only 30%  got covid-19 from taking the vaccine. So the vaccine does not prevent males getting covid-19 as well as taking the placebo.  

## Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
malesfish <- fisher.test(table1)
malesfish
```
  
For this chi-squared test, shows that the males who took the vaccine are more likely to get covid-19. The p-value for this shows a 4.631e-14, which is not a good number to see. The lower the p-value have the better results. Covid-19 does show that the illness can also be based off gender, since each body is different and reacts differently from covid-19. 

## Effectiveness Confidence Interval

```{r}
effect <- malesfish$p.value/(malesfish$p.value +1)
effectivenessguy = malesfish$estimate/(malesfish$estimate +1)*100
print(effectivenessguy)
```
```{r}
low=(malesfish$conf.int[1]/(malesfish$conf.int[1] +1))*100
hi=(malesfish$conf.int[2]/(malesfish$conf.int[2] +1))*100
options(digits=2)
```
  
For the odds ratio it shows that it is at a 69.15981, which is a high number that the odds of people taking the vaccine are getting effected by covid-19. 


## Females

```{r}
dd2 <- females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
This plot shows that the vaccine protected more females from covid-19 than the placebo. There was a higher percentage of females who took the placebo but still ended up getting infected. There was 60% of females who got infected by taking placebo and about 39% from the vaccine. 

## Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```
  
About 60 females got infected by covid-19 from taking the placebo while 40 females got infected by covid-19 from the vaccine. Out of 100 females, 50 on both sides were okay from taking the placebo and the vaccine. For the 2nd group, there were more people that did not get covid-19 from either the placebo or vaccine. 

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
femalesfish <- fisher.test(table1)
femalesfish
```
  
Since the p-value (3e-04) is less than the conventional significance level of .05, we can reject the null hypothesis and conclude that there is a significant association between the infection status and the treatment.

## Efectiveness Confidence Interval

```{r}
effect <- femalesfish$p.value/(femalesfish$p.value +1)
effectivenessfemale = femalesfish$estimate/(femalesfish$estimate +1)*100
print(effectivenessfemale)
```
```{r}
low=(femalesfish$conf.int[1]/(femalesfish$conf.int[1] +1))*100
hi=(femalesfish$conf.int[2]/(femalesfish$conf.int[2] +1))*100
options(digits=2)
```



## LGBTQ



## Druggies


# Overall Results and Conclusions